---
title: "`r params$title`"
author: "Christophe Bruley"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    code_folding: "hide"
  word_document: 
    fig_height: 5
  html_document:
    number_sections: yes
    toc: yes
    code_folding: "hide"
params:
  title: "(ABRF)"
  input_directory: 'data/ABRF/20200123_peakel/spec_sumion_sum'
  ions_sheet_index: 5
  proteins_sheet_index: 4
  peptide_sheet_index: 7
  flags_accession_regexp: !r list(background = '_YEAS|_yeas|_SCV?|_SACPS')
  abundance_column_regexp: '^abundance'
  dapar_normalize: TRUE
  dapar_imputation: 'quantile'
  dapar_quantile_qval: 0.025
  dapar_test_method: 'Limma'

---

```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
library(reshape2)
library(matrixcalc)
library(openxlsx)
library(rmdformats)

knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)

```

```{r read_data}

source('R/commons_functions.R')
source('R/file-utils.R')

file = list.files(path = params$input_directory, pattern = ".*\\.xlsx")[1]
all.proteins = NULL
ions.summary = NULL
proteins.summary = NULL
peptides.summary = NULL

```


# Parameters 

Parameters used to generate this report : 
`r knitr::kable(knit_params_as_dataframe(params))`

Detected data files to process: `r paste0(file)`

# Comparison `r file`


```{r child="dataset_summary.Rmd" }
```


# Differential Analysis

The differential analysis reads the proteins data again to overcome the error due to a spiked protein identified among contaminants. **Warning** This means that the error is not taken into account in the previous sections of this document.

```{r differential_analysis}

filename = paste(params$input_directory, file ,sep = '/')
proline.proteins = openxlsx::read.xlsx(filename, sheet = params$proteins_sheet_index, colNames = TRUE, startRow = 0)
proline.proteins = dplyr::select(proline.proteins, matches(paste0('accession|', params$abundance_column_regexp)))

proline.proteins[which(proline.proteins$accession == "#C#P02769"), ]$accession = "P02769"
proline.proteins = proline.proteins %>% filter(!grepl("^#C#", accession))
proline.proteins = flag_by_accession(proline.proteins, params$background_accession_regexp)

conditions = c("1", "2", "3", "4")
columns = group_columns_by_conditions(proline.proteins, conditions , params$abundance_column_regexp)
column.other.indexes = which(!(colnames(proline.proteins) %in% list.flatten(columns)))


pairwise.conditions = combn(c("1", "2", "3", "4"), 2)
differential.proteins = NULL
all.proteins = NULL

for (i in 1:ncol(pairwise.conditions)) {

  condition = paste0(pairwise.conditions[,i],collapse='-')
  columns.indexes = which(colnames(proline.proteins) %in% list.flatten(columns[pairwise.conditions[,i]]))
  
  subset.proline.proteins = proline.proteins[,c(column.other.indexes, columns.indexes)]
  subset.proline.proteins = dapar_differential_protein_analysis(subset.proline.proteins, 
                                                                condition, 
                                                                params$abundance_column_regexp,
                                                                params$dapar_normalize, 
                                                                params$dapar_imputation, 
                                                                params$dapar_quantile_qval, 
                                                                params$dapar_test_method)

  ggplot(subset.proline.proteins, aes(x = logFC, y = -log10(pvalue))) +
  geom_point(aes(color = factor(flag)), size = 2.0, shape = 16, alpha = .6) +
  theme(legend.position = "bottom")

  all.proteins =rbind(all.proteins, subset.proline.proteins)
  
  subset.proline.proteins = dplyr::filter(subset.proline.proteins, !is.na(logFC))
  
  # subset.proline.proteins$adjusted_pvalue = cp4p::adjust.p(subset.proline.proteins$pvalue)$adjp$adjusted.p
  # subset.proline.proteins = dplyr::arrange(subset.proline.proteins, pvalue)
  # index = max(which(subset.proline.proteins$adjusted_pvalue < 0.01))
  # differential.proteins = rbind(differential.proteins, dplyr::slice(subset.proline.proteins, 1:index))

  differential.proteins = rbind(differential.proteins, subset.proline.proteins)
  
}

```


## Intensities distribution

```{r }
differential.ref = c("P02769","UTR6_YEAST","SFG2_YEAST","ZRT4_YEAST","ISCB_YEAST","VAC2_YEAST")
all.proteins[is.element(all.proteins$accession, differential.ref), "flag"] = "spkiked"
all.proteins = all.proteins %>% mutate(C1_mean = rowMeans(select(., starts_with("C1")), na.rm = TRUE))

ggplot(all.proteins) + geom_point(aes(x = log10(C1_mean), y = logFC, color = flag), shape = 16, alpha = .6)

``` 

## Declared as differential (1% FDR)

```{r results='asis'}

  differential.proteins$adjusted_pvalue = cp4p::adjust.p(differential.proteins$pvalue)$adjp$adjusted.p
  differential.proteins = dplyr::arrange(differential.proteins, pvalue)
  index = max(which(differential.proteins$adjusted_pvalue < 0.01))

  differential.proteins = dplyr::slice(differential.proteins, 1:index)

  knitr::kable(flags_summary(differential.proteins))
  knitr::kable(dplyr::select(differential.proteins, -c(2,3,4)))

  differential.ref = c("P02769","UTR6_YEAST","SFG2_YEAST","ZRT4_YEAST","ISCB_YEAST","VAC2_YEAST")
  differential.true = dplyr::filter(differential.proteins, accession %in% differential.ref)

  cat('True spiked proteins declared as differential')

  knitr::kable(flags_summary(differential.true))


```

## Observed vs Expected ratios


```{r }
expectedFC = data.frame(accession = differential.ref, 
                        "1-2" = c(-4.196397213 ,5.022367813 ,-2.906890596 ,5.64385619 ,-1.874469118 ,-0.2410081),
                        "1-3" = c(-0.137503524 ,4.781359714 ,2.115477217 ,0.137503524 ,-4.781359714 ,-2.115477217),
                        "1-4" = c(5.506352666 ,2.906890596 ,1.874469118 ,-4.058893689 ,0.2410081 ,-5.022367813),
                        "2-3" = c(4.058893689 ,-0.2410081 ,5.022367813 ,-5.506352666 ,-2.906890596 ,-1.874469118),
                        "2-4" = c(9.702749879 ,-2.115477217 ,4.781359714 ,-9.702749879 ,2.115477217 ,-4.781359714),
                        "3-4" = c(5.64385619 ,-1.874469118 ,-0.2410081 ,-4.196397213 ,5.022367813 ,-2.906890596))
expectedFC = melt(expectedFC, id = "accession")
expectedFC$variable = gsub('X', '', expectedFC$variable)
expectedFC$variable = gsub('\\.', '-', expectedFC$variable)
colnames(expectedFC) = c("accession", "dataset", "expectedFC")

spiked.proteins = all.proteins[is.element(all.proteins$accession, differential.ref),]
spiked.proteins = left_join(spiked.proteins, expectedFC, by = c("accession", "dataset"))
spiked.differential = data.frame(accession = differential.true$accession, dataset = differential.true$dataset, differential = TRUE)

spiked.proteins = left_join(spiked.proteins, spiked.differential, by = c("accession", "dataset"))
spiked.proteins[which(is.na(spiked.proteins$differential)), "differential"] = FALSE

ggplot(spiked.proteins) + geom_point(aes(x = abs(expectedFC), y = abs(logFC), color = differential)) + 
  geom_abline(intercept =0 , slope = 1, linetype = "12") 

```