
## Ions analysis 

\newline

```{r echo=FALSE}
filename = paste(params$input_directory, file ,sep = '/')

sheet.names = openxlsx::getSheetNames(filename)

proline.ions = openxlsx::read.xlsx(filename, sheet = params$ions_sheet_index, colNames = TRUE, startRow = 0)
proline.ions = cleanup_ions_dataset(proline.ions, params$abundance_column_regexp)
proline.ions = flag_by_accession(proline.ions, params$flags_accession_regexp)
proline.ions = replace_zero_by_na(proline.ions, params$abundance_column_regexp)

if (is.element("background", names(params$flags_accession_regexp))) {
  background_accession_regexp = params$flags_accession_regexp[["background"]]
} else {
  background_accession_regexp = NULL
}

background.ions = ions_analysis(proline.ions, background_accession_regexp, params$abundance_column_regexp)

```

* Analysis parameters : 
    + The analysed file name : **`r filename`**
    + File date: **`r file.info(filename)$ctime`**
    + Ions sheet used : **`r sheet.names[[params$ions_sheet_index]]`** 
    + Proteins sheet used : **`r sheet.names[[params$proteins_sheet_index]]`** 
    + Number of runs: **`r ncol(background.ions$abundances) `**
    + Number of ions : **`r nrow(proline.ions)`**

### Ions Summary

Summary of identified and quantified (fully or partially) ions, grouped by matching accessions categories (see `flags_accession_regexp` parameter).

```{r results='asis'}

ions_summary = quantified_summary(proline.ions, params$abundance_column_regexp)
knitr::kable(ions_summary$counts)
knitr::kable(ions_summary$rates) 

## add the file summary to the global one
ions_summary = ions_summary$counts
ions_summary$file = file
ions.summary = rbind(ions.summary, ions_summary)

```
   
### Ions Missing values 

Number of ions matching a "background" (see `flags_accession_regexp` parameter, if "background" category is undefined, takes all ions into account) protein for which an abundance value was defined in more than *n* samples.

```{r, fig.align='center', fig.width=8 }

ggplot(background.ions$missing, aes(runs_count, y = value, color = variable)) + 
  geom_point(aes(y = with_cross, col = 'with cross-assignment'), size = 3) +
  geom_point(aes(y = without_cross, col = 'without cross-assignment'), size = 3) +
  ylim(0, NA) + labs(title="Background ions quantifed in at least x runs ", x ="# runs", y = "# ions") +
  theme(legend.position = "bottom")

```

### Ions CVs distribution 


ions CVs distribution with/without cross-assigned values. 

```{r results='asis'}

cvs_summary = rbind( nocross = summary(na.omit(background.ions$CVs$without_cross)),
                     cross = summary(na.omit(background.ions$CVs$with_cross)))
knitr::kable(cvs_summary)

```


ions CVs density plot.

```{r fig.width=8, fig.align='center'}
#
# Plot CV distribution
#
cvs = reshape2::melt(background.ions$CVs)
median.cross = median(background.ions$CVs$with_cross, na.rm = TRUE)
median.nocross = median(background.ions$CVs$without_cross, na.rm = TRUE)

p = ggplot(data = cvs, aes(x = value, fill = variable)) + geom_density(aes(y=..count..), alpha = 0.2) + 
  geom_vline(xintercept = median.cross, col = "darkorange3") + 
  geom_vline(xintercept = median.nocross, col = "blue") +
  scale_x_continuous(breaks = seq(0, 60, by = 2), limits = c(0,60)) + 
  labs( x = "Coefficient of Variation (%)", y = "Density") 

p.data <- ggplot_build(p)$data[[1]]
p.groups <- lapply(split(p.data, f = p.data$group), function(df){
  df[which.max(df$count), ]
})

p + 
  annotate(geom= "label", label = sprintf("%1.1f", median.cross), x = median.cross,y = p.groups[[1]]$count/4 , size = 3, col = "darkorange3") + 
  annotate(geom= "label", label = sprintf("%1.1f", median.nocross), x = median.nocross, y = p.groups[[1]]$count/4+100, size = 3, col = "blue") +
  theme(legend.position = "bottom")

```


## Peptides analysis 


\newline

```{r echo=FALSE}

apply_peptides_analysis = is.element("peptides_sheet_index", names(params)) 

if(apply_peptides_analysis) {
  proline.peptides = openxlsx::read.xlsx(
    filename,
    sheet = params$peptides_sheet_index,
    colNames = TRUE,
    startRow = 0
  )
  proline.peptides = cleanup_peptides_dataset(proline.peptides, params$abundance_column_regexp)
  proline.peptides = flag_by_accession(proline.peptides, params$flags_accession_regexp)
  proline.peptides = replace_zero_by_na(proline.peptides, params$abundance_column_regexp)

  
  if (is.element("background", names(params$flags_accession_regexp))) {
    background_accession_regexp = params$flags_accession_regexp[["background"]]
  } else {
    background_accession_regexp = NULL
  }
  
  background.peptides = ions_analysis(proline.peptides,
                                      background_accession_regexp,
                                      params$abundance_column_regexp)
}

```


### Peptides Summary

Summary of identified and quantified (fully or partially) peptides, grouped by matching accessions categories (see `flags_accession_regexp` parameter).

```{r eval=apply_peptides_analysis}

peptides_summary = quantified_summary(proline.peptides, params$abundance_column_regexp)

## add the file summary to the global one
peptides_count_summary = peptides_summary$counts
peptides_count_summary$file = file
peptides.summary = rbind(peptides.summary, peptides_count_summary)

```


```{r results='asis', eval=apply_peptides_analysis}

knitr::kable(peptides_summary$counts)
knitr::kable(peptides_summary$rates)

```

### peptides Missing values 

Number of peptides matching a "background" (see `flags_accession_regexp` parameter, if "background" category is undefined, takes all peptides into account) protein for which an abundance value was defined in more than *n* samples.

```{r, fig.align='center', fig.width=8, eval=apply_peptides_analysis}

ggplot(background.peptides$missing,aes(runs_count, y = value, color = variable)) +
  geom_point(aes(y = with_cross, col = 'with cross-assignment'), size = 3) +
  geom_point(aes(y = without_cross, col = 'without cross-assignment'), size = 3) +
  ylim(0, NA) + labs(title = "Background peptides quantifed in at least x runs ", x = "# runs", y = "# peptides") +
  theme(legend.position = "bottom")

```

### peptides CVs distribution 


peptides CVs distribution with/without cross-assigned values. 

```{r results='asis', eval=apply_peptides_analysis}

cvs_summary = rbind( nocross = summary(na.omit(background.peptides$CVs$without_cross)),
                     cross = summary(na.omit(background.peptides$CVs$with_cross)))
knitr::kable(cvs_summary)

```


peptides CVs density plot.

```{r fig.width=8, fig.align='center', eval=apply_peptides_analysis}

#
# Plot CV distribution
#
cvs = reshape2::melt(background.peptides$CVs)
median.cross = median(background.peptides$CVs$with_cross, na.rm = TRUE)
median.nocross = median(background.peptides$CVs$without_cross, na.rm = TRUE)

p = ggplot(data = cvs, aes(x = value, fill = variable)) + geom_density(aes(y=..count..), alpha = 0.2) +
  geom_vline(xintercept = median.cross, col = "darkorange3") +
  geom_vline(xintercept = median.nocross, col = "blue") +
  scale_x_continuous(breaks = seq(0, 60, by = 2), limits = c(0, 60)) +
  labs(x = "Coefficient of Variation (%)", y = "Density") 

p.data <- ggplot_build(p)$data[[1]]
p.groups <- lapply(split(p.data, f = p.data$group), function(df){
  df[which.max(df$count), ]
})

p +
  annotate(geom = "label", label = sprintf("%1.1f", median.cross), x = median.cross, y = p.groups[[1]]$count/4,size = 3,col = "darkorange3") +
  annotate(geom = "label", label = sprintf("%1.1f", median.nocross),x = median.nocross,y = p.groups[[1]]$count/4 + 100,size = 3,col = "blue") +
  theme(legend.position = "bottom")

```



## Proteins analysis

The data are read from the proteins sheet of the analyzed Excel file. **Contaminant are filtered**  

```{r }

proline.proteins = openxlsx::read.xlsx(filename, sheet = params$proteins_sheet_index, colNames = TRUE, startRow = 0)
proline.proteins = dplyr::select(proline.proteins, matches(paste0('accession|', params$abundance_column_regexp)))
proline.proteins = proline.proteins %>% filter(!grepl("^#C#", accession))
proline.proteins = flag_by_accession(proline.proteins, params$flags_accession_regexp)
proline.proteins = replace_zero_by_na(proline.proteins, params$abundance_column_regexp)

```

### Proteins Summary

```{r results='asis'}  

proteins_summary = quantified_summary(proline.proteins, params$abundance_column_regexp)
knitr::kable(proteins_summary$counts)
knitr::kable(proteins_summary$rates)

proteins_summary = proteins_summary$counts
proteins_summary$file = file
proteins.summary = rbind(proteins.summary, proteins_summary)

```


### Proteins CVs distribution

```{r }

background.proteins = abundance_values_analysis(proline.proteins, background_accession_regexp, params$abundance_column_regexp)

median.cv = median(background.proteins$CVs, na.rm = TRUE)

p = ggplot(data = data.frame(cv=background.proteins$CVs), aes(x = cv)) + geom_density(aes(y=..count..), alpha = 0.2, fill = "darkorange3") + 
  geom_vline(xintercept = median.cv, col = "darkorange3") +
  scale_x_continuous(breaks = seq(0, 150, by = 5), limits = c(0,150)) + 
  labs( x = "Coefficient of Variation (%)", y = "Density") 

p.data <- ggplot_build(p)$data[[1]]
p.groups <- lapply(split(p.data, f = p.data$group), function(df){
  df[which.max(df$count), ]
})

p +
  annotate(geom= "label", label = sprintf("%1.1f", median.cv), x = median.cv, y = p.groups[[1]]$count/4, size = 3, col = "darkorange3") +
  theme(legend.position = "bottom")

```


```{r }
condition = gsub(".*_(.*)\\.xlsx", "\\1", file)
if (condition == file) {
  condition = gsub("(.*)\\.xlsx", "\\1", file)
}
apply_diff_analysis = grepl('-', condition) && is.element("dapar_test_method", names(params)) 

```

```{r results='asis', eval=apply_diff_analysis} 

  cat('### Proteins differential analysis')

```

  
```{r fig.width=8, eval=apply_diff_analysis}

r = dapar_differential_protein_analysis(proline.proteins, 
                                      condition, 
                                      params$abundance_column_regexp, 
                                      params$dapar_normalize, 
                                      params$dapar_imputation, 
                                      params$dapar_quantile_qval, 
                                      params$dapar_test_method)

all.proteins = rbind(all.proteins, r)

ggplot(r, aes(x = logFC, y = -log10(pvalue))) +
  geom_point(aes(color = factor(flag)), size = 3.5, shape = 16, alpha = .6) +
  scale_x_continuous(breaks = seq(-9, 9, by = 1), limits = c(-9,9)) +
  theme(legend.position = "bottom")


```

Intensities distribution 

```{r eval=apply_diff_analysis}

ma.proteins = all.proteins %>% mutate(C1_mean = rowMeans(select(., starts_with("C1")), na.rm = TRUE))
ggplot(ma.proteins) + geom_point(aes(x = log10(C1_mean), y = logFC, color = flag), shape = 16, alpha = .6)

```