---
title: "`r params$title`"
author: "Christophe Bruley"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    code_folding: "hide"
  word_document: 
    fig_height: 5
  html_document:
    number_sections: yes
    toc: yes
    code_folding: "hide"
params:
  title: '(PME12)'
  input_directory: 'data/PME12/20191216/spec_bestion_sum'
  ions_sheet_index: 5
  proteins_sheet_index: 4
  flags_accession_regexp: !r list(human = '_HUMAN', yeast = '_YEAS|_yeas|_SCV?|_SACPS', ecoli = '_ECO')
  abundance_column_regexp: '^abundance'
  dapar_normalize: TRUE
  dapar_imputation: 'standard'
  dapar_quantile_qval: 0.025
  dapar_test_method: 'Limma'

---

```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
library(reshape2)
library(matrixcalc)
library(openxlsx)
library(rmdformats)

knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)

```

```{r read_data }

source('R/commons_functions.R')
source('R/file-utils.R')

file = list.files(path = params$input_directory, pattern = ".*\\.xlsx")[1]
all.proteins = NULL
ions.summary = NULL
proteins.summary = NULL

```

# Parameters 

Parameters used to generate this report : 
`r knitr::kable(knit_params_as_dataframe(params))`

Detected data files to process: `r paste0(file)`

# Comparison `r file`


```{r child="dataset_summary.Rmd" }
```


# Differential Analysis

```{r differential_analysis }
condition = "A-B"

proline.proteins = dapar_differential_protein_analysis(proline.proteins, 
                                      condition, 
                                      params$abundance_column_regexp, 
                                      params$dapar_normalize, 
                                      params$dapar_imputation, 
                                      params$dapar_quantile_qval, 
                                      params$dapar_test_method)


ggplot(proline.proteins, aes(x = logFC, y = -log10(pvalue))) + 
  scale_x_continuous(breaks = seq(-9, 9, by = 1), limits = c(-9,9)) +
  geom_point(aes(color = factor(flag)), size = 3.5, shape = 16, alpha = .6) +
  theme(legend.position = "bottom")

```



## Intensities distribution

```{r }
ma.proteins = proline.proteins %>% mutate(C1_mean = rowMeans(select(., starts_with("C2")), na.rm = TRUE))
ggplot(ma.proteins) + geom_point(aes(x = log10(C1_mean), y = logFC, color = flag), shape = 16, alpha = .6)

``` 

## Declared as differential (1% FDR)

```{r }
proline.proteins = dplyr::filter(proline.proteins, !is.na(logFC))
proline.proteins$adjusted_pvalue = cp4p::adjust.p(proline.proteins$pvalue)$adjp$adjusted.p
proline.proteins = dplyr::arrange(proline.proteins, pvalue)
index = max(which(proline.proteins$adjusted_pvalue < 0.01))
differential.proteins = dplyr::slice(proline.proteins, 1:index)

differential_summary = flags_summary(differential.proteins)
```


```{r results='asis'}
knitr::kable(differential_summary)
```

## Declared as differential (1% FDR, abs(logFC) > 0.8)

```{r results='asis'}

differential.proteins = differential.proteins[abs(differential.proteins$logFC) >= 0.8, ]
differential_summary = flags_summary(differential.proteins)

knitr::kable(differential_summary)

proteins.summary = proteins.summary[which(proteins.summary$name=="quantified"), ]

```

* % of correctly identified as differential:
  + Human : **`r 100*differential_summary$human/proteins.summary$human`%**
  + Yeast: **`r 100*differential_summary$yeast/proteins.summary$yeast`%**
  + EColi: **`r 100*differential_summary$ecoli/proteins.summary$ecoli`%**



