---
title: "`r params$title`"
author: "Christophe Bruley"
date: "`r Sys.time()`"
output: 
  rmdformats::readthedown:
    highlight: kate
    code_folding: "hide"
  html_document:
    number_sections: yes
    toc: yes
    code_folding: "hide"
  word_document: 
    fig_height: 5
params:
  title: '(UPS_4Concentrations)'
  input_directory: 'data/UPS_4Concentrations/2019_v1.6papier/spec_bestion_sum'
  proteins_sheet_index: 4
  peptides_sheet_index: 7
  ions_sheet_index: 5
  flags_accession_regexp: !r list(ups='HUMAN_UPS', background='YEAS|yeas')
  abundance_column_regexp: '^abundance'
  proteins_output_file: 'Proteins.xlsx'
  dapar_normalize: TRUE
  dapar_imputation: 'quantile'
  dapar_quantile_qval: 0.025
  dapar_test_method: 'Limma'

---

```{r setup, include=FALSE, cache = FALSE}

library(dplyr)
library(ggplot2)
library(reshape2)
library(matrixcalc)
library(openxlsx)
library(rmdformats)

knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)

```


```{r read_data}

source('R/commons_functions.R')
source('R/file-utils.R')

files = list.files(path = params$input_directory, pattern = ".*\\.xlsx")

all.proteins = NULL
ions.summary = NULL
proteins.summary = NULL
peptides.summary = NULL

```

*Warning*\
The following report read Proline xlsx exported files contained in the specified directory \
The naming convention for UPS_4Concentration report is supposed to be \<file-name-prefix\>_**ConditionA-ConditionB**.xlsx \


# Report Parameters 

Parameters used to generate this report : 
`r knitr::kable(knit_params_as_dataframe(params))`

Detected data files to process: 

```{r results='asis'} 
cat(paste('*', files), sep = '\n')
```

# Comparison `r files[1]`

```{r first file}
file = files[1]
```

```{r child="dataset_summary.Rmd" }
```

# Comparison `r files[2]`

```{r second file}
file = files[2]
```

```{r child="dataset_summary.Rmd" }
```

# Comparison `r files[3]`

```{r third file}
file = files[3]
```

```{r child="dataset_summary.Rmd" }
```

# Combined analysis

## combined Ions

```{r results='asis'}
knitr::kable(ions.summary)
```


## Combined Proteins

Without contaminants and before differential analysis:

```{r results='asis'}
knitr::kable(proteins.summary)
```

\newline

```{r roc_analysis, fig.align='center', fig.width=8}

roc = compute_roc_curve(all.proteins, "ups")

roc$label = params$input_directory
pvalue_marks = c(0.05, 0.01, 0.001)
points = NULL

for (mark in pvalue_marks) {
  points = rbind(points, roc[which.min(abs(roc$pvalue - mark)),])
}
point_label = paste(pvalue_marks)   # paste(round(points$pvalue,5))

ggplot(data = roc, aes(x = fdp, y = tpr, color = label)) + 
  geom_path(size = 1.3) +
  geom_point() +
  scale_x_continuous(breaks = round(seq(min(roc$fdp), max(roc$fdp), by = 5), 1),
                     minor_breaks = round(seq(min(roc$fdp), max(roc$fdp), by = 1), 1)) +
  scale_y_continuous(breaks = round(seq(0, 100, by = 10), 1)) +
  geom_point(data = points, aes(x = points$fdp, y = points$tpr), size = 2, color = "black") +
  geom_text(data = points, aes(x = points$fdp, y = points$tpr, label = point_label),
            color = "black",
            hjust = 0,
            nudge_x = 2.0) +
  theme(legend.position = "bottom")

sheets_list = list("proteins" = all.proteins, "roc" = roc)
openxlsx::write.xlsx(sheets_list, file = params$proteins_output_file)


```

```{r thresholds} 

all.proteins$adjusted_pvalue = cp4p::adjust.p(all.proteins$pvalue)$adjp$adjusted.p
all.proteins = dplyr::arrange(all.proteins, pvalue)

qvalue_marks = c(0.05, 0.01, 0.001)
points = NULL

for(qvalue_threshold in qvalue_marks) {
 index = max(which(all.proteins$adjusted_pvalue < qvalue_threshold))
 counts = all.proteins %>% slice(1:index) %>% count(flag)
 counts$qvalue = qvalue_threshold
 counts = reshape2::dcast(reshape2::melt(counts, c("flag","qvalue")), qvalue~flag)

 points = rbind(points, counts)
}

# search for the 0% FDP point

index = min(which(all.proteins$flag == "background"))
counts = all.proteins %>% slice(1:(index-1)) %>% count(flag)
counts$qvalue = 0
counts = reshape2::dcast(reshape2::melt(counts, c("flag","qvalue")), qvalue~flag)
counts$background = 0

points = rbind(points, counts)
points$tpr = 100*points$ups/144
points$fdp = 100*points$background / (points$ups + points$background)

knitr::kable(points)

```
